---
title: "Research Question 2"
output: html_document
author: Anran Huang
date: 10/05/2021
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading packages and the dataset, message=FALSE, warning=FALSE}
# check package availability
list.of.packages <- c("ggplot2", "epiDisplay", "ROSE", "caret" ,"effects", 
                      "GGally", "compiler", "parallel", "devtools", "texreg", 
                      "texreghelpr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

#load packages
library(nlme)
library(ggplot2)
library(lme4)
library(nlme)
library(dplyr)
library(caret)
library(tidyr)
require(GGally)
require(reshape2)
require(compiler)
require(parallel)
require(boot)
require(lattice)
library(devtools)
library(texreg)
install_github("SarBearSchwartz/texreghelpr")
library(texreghelpr)


icudata <- read.csv('1_input/input_rq2.csv')
paste(nrow(icudata), 'rows of data')
paste(length(unique(icudata$icustay_id)), 'unique ICU stays in the dataset')
paste(length(unique(icudata$hadm_id)), 'unique hospital admissions in the dataset')
paste(length(unique(icudata$subject_id)), 'patients in the dataset')
```

## Data Preparation

```{r data wrangling}
colnames(icudata)[8] <- 'ICU_mortality'
data_long <- icudata %>% 
  group_by(hadm_id) %>% 
  mutate(icustay_idx = row_number()) %>% ungroup() %>%
  dplyr::mutate(hadm_id = factor(hadm_id),
                subject_id = factor(subject_id),
                age_bins = factor(age_bins, 
                                  order = TRUE, 
                                  levels = c("under44", "45-54", "55-64", "65-74", "over75")),
                icu_adm_weekend = as.logical.factor((icu_adm_weekend),
                                                    levels = c(FALSE, TRUE),
                                                    labels = c("Weekday", "Weekend")),
                ICU_mortality = as.factor(as.integer(as.integer(ICU_mortality)-1)),
                ICU_mortality_int = as.integer(ICU_mortality),
                icu_adm_weekend = factor((icu_adm_weekend),
                                         levels = c(FALSE, TRUE),
                                         labels = c("Weekday", "Weekend"))) %>%
  select(-c(icustay_id)) %>%
  dplyr::arrange(hadm_id, icustay_idx)
str(data_long)
```

## Data Visualisation

The longitudinal data set is seriously imbalanced. 93% of the cases are negative and only 7% are positive (i.e. ICU mortality).

```{r icu_adm_weekend as fixed effect}
round(prop.table(table(data_long$ICU_mortality)),3)
```

```{r why icu_adm_weekend as fixed effect}
cross.tab <- xtabs(~icu_adm_weekend+ICU_mortality, data=data_long)
icu.crosstab <- t(cross.tab[2:1,2:1])
addmargins(icu.crosstab) #cross tab
round(prop.table(icu.crosstab,1),3) # row perc
round(prop.table(icu.crosstab,2),3) # column perc
round(prop.table(icu.crosstab),3) # cell perc
mosaicplot( t(icu.crosstab), col = c("firebrick", "goldenrod1"), cex.axis = 1, sub = "icu_adm_weekend", ylab = "Relative frequency", main = "")
```
```{r odds ratio and risk ratio}
if (!require("epitools")) {install.packages("epitools", dependencies = TRUE, repos="http://cran.rstudio.com/")}
library(epitools)
cat("odds ratio: \n")
oddsratio(icu.crosstab, method = "wald")$measure[-1,]
cat("risk ratio: \n")
riskratio(icu.crosstab, method = "wald", rev = "both")$measure[-1,]
```
```{r categorical viz}
par(mfrow = c(3,3), mar=c(0.5, 0.5, 0.2, 0.2),
     mai=c(.5,.6,.2,.2))
loop.vector <- c("icustay_idx", "age_bins", "icu_adm_weekend", "re_adm_in30d", "insurance", "admission_type")

for (i in loop.vector) {
  cat(paste("Frequency Table for", i, '\n'))
  print(epiDisplay::tab1(data_long[,c(i)], sort.group ='decreasing', cum.percent = FALSE, graph=TRUE, cex=0.8, bar.values='percent', main=paste("Bar Chart for",i)))
}

```

```{r continuous viz}
print(data_long %>% group_by(ICU_mortality) %>% 
        summarise(median_los = median(los), 
                  qntl_25 = quantile(los, 0.25),
                  qntl_75 = quantile(los, 0.75),
                  ))
data_long %>%
  ggplot(aes(x=ICU_mortality, y=los, fill=ICU_mortality)) +
  geom_violin() + 
  ylab("length of ICU stay in days") +
  geom_boxplot(width=0.1, color="orange", fill='white', alpha=0.5)+ 
  ylim(0, 5)+
  #scale_y_continuous(limits = quantile(data_long$los, c(0, 0.9)))+
  ggtitle("Violin and Box Plot for ICU Lenth of Stay vs ICU Mortality") +
  xlab("")
```

## Chi-squared contingency test to test association between ICU weekend admission and ICU mortality rate

```{r chi-sq test}
icu_ChiTest <- chisq.test(data_long$ICU_mortality, data_long$icu_adm_weekend, correct = FALSE)
icu_ChiTest
addmargins(icu_ChiTest$expected)
source("http://www.stat.wisc.edu/~st571-1/gtest.R")
g.test(icu.crosstab)
```

## Under-sampling for Label Imbalance 

```{r undersampling}
data_long$age_group = as.integer(data_long$age_bins)
sliced = data_long[,c("ICU_mortality", "hadm_id", "subject_id", "los", "age_group", "icu_adm_weekend", "re_adm_in30d", "insurance", "admission_type")]
under_sample <- ROSE::ovun.sample(data=sliced, ICU_mortality ~ .,method = "under", N = 4323*2, seed = 1)$data
```

```{r age group as fixed effect}
xtabs(~age_group+ICU_mortality, data=data_long)
```

## Simple Generalised Logistic Regression Model (SLR)

```{r glm simple}
model_slm0 <- glm(ICU_mortality ~ icu_adm_weekend, data = under_sample, family = binomial)
# Summarize the model
summary(model_slm0)
model_slm0$coefficients
confint(model_slm0)
probabilities0 <-model_slm0 %>% predict(data_long, type = "response")
predicted.classes0 <- ifelse(probabilities0 > 0.5, 1, 0)
cat("model accuracy on predicting ICU mortality:", mean(predicted.classes0 == data_long$ICU_mortality))
# use caret and compute a confusion matrix
caret::confusionMatrix(data = as.factor(predicted.classes0), reference = data_long$ICU_mortality)
```

## Multilevel Linear Model for Binary Outcome (MLM)

We are interested in investigating whether weekend ICU admission increases the risk of ICU mortality. 

- RI: outcome may change for each hospital admission `icu_adm_weekend`;

Random slope models are generally not recommended for binomial GLMMs.

By comparing the log-likelihood of single-level intercept-only model and a variance-components (VC) model, it does appear the VC model with a random intercept is better than a single-level model. As expected, the confidence interval is wider for the VC model as the single-level is often over-confident. VPC of 0.6 at hospital admission level, which is not too large nor close to zero. It suggests that there is an impact on `ICU_mortality` by `hadm_id`, i.e. risk of `ICU_mortality` varies among `hadm_id`s.

```{r variance-components vs single-level}
under_sample$ICU_mortality = as.numeric(under_sample$ICU_mortality)
mod.single <- lm(ICU_mortality ~ 1, data=under_sample)
mod.vc.hosp <- lmer(ICU_mortality ~ (1|hadm_id), data=under_sample, REML=FALSE)
mod.vc.pt <- lmer(ICU_mortality ~ (1|subject_id), data=under_sample, REML=FALSE)

VPC.hosp <- VarCorr(mod.vc.hosp) %>% as_tibble() %>% mutate(icc=vcov/sum(vcov)) %>% select(grp, icc)
VPC.pt <- VarCorr(mod.vc.pt) %>% as_tibble() %>% mutate(icc=vcov/sum(vcov)) %>% select(grp, icc)

print(VPC.hosp[0:1,])
print(VPC.pt[0:1,])

logLik(mod.single)
logLik(mod.vc.hosp)
logLik(mod.vc.pt)

confint(mod.single)
confint(mod.vc.hosp)
confint(mod.vc.pt)
```

```{r RI models}
under_sample$ICU_mortality = as.factor(under_sample$ICU_mortality)
# random intercept model 1
mod1 <- glmer(ICU_mortality ~ icu_adm_weekend + (1|hadm_id), data=under_sample, family = binomial(link = "logit"))

# additional model considering age_group
mod2 <- glmer(ICU_mortality ~ icu_adm_weekend + age_group + (1|hadm_id), data=under_sample, family = binomial(link = "logit"))

# additional model considering age_group * icu_adm_weekend
mod3 <- glmer(ICU_mortality ~ icu_adm_weekend*age_group + (1|hadm_id), data=under_sample, family = binomial(link = "logit"))
```

```{r mlm 1 summary}
confint(mod1)
summary(mod1)
logLik(mod1)
logLik(model_slm0)
plot(effects::allEffects(mod1))
```
```{r compare mlms with or without age_group}
texreg::knitreg(list(mod1, mod2),
                custom.model.names = c("without age group", "with age group"),
                single.row = TRUE,
                caption = "MLM Parameter Estimates: Inclusion of Age Group at Hospital Admission (SE and p-values)")
plot(effects::allEffects(mod2))
```
```{r odds ratio scale comparison for mlms with or without age_group}
texreg::knitreg(list(extract_glmer_exp(mod1), 
                     extract_glmer_exp(mod2)),
                custom.model.names = c("without age group", "with age group"),
                ci.test = 1,
                single.row = TRUE,
                caption = "MLM Parameter Estimates: Inclusion of Age Group at Hospital Admission (95% CI's)")
```
